Using CopyKAT to idenfity tumor cells in Ewing sarcoma samples

### 2024-05-13 analyses/ewings-cell-type/03-copykat.Rmd
plot each chr and remove UMAPs

# filter to only have cnvs found in Ewings to keep things a bit smaller when joining
ewings_chrs <- c(1, 8, 12, 16)

ewings_cnv_df <- full_cnv_df |> 
  dplyr::filter(chromosome_name %in% ewings_chrs)

# join with 
cnv_df <- cnv_df |> 
  dplyr::left_join(ewings_cnv_df, by = c("barcodes", "reference_used"))
```

```{r}
# create faceted UMAPs showing estimation of CNV detection across each chr of interest 
ewings_chrs |> 
  purrr::map(\(chr){
    plot_df <- cnv_df |> 
      dplyr::filter(chromosome_name == chr)

    ggplot(plot_df, aes(x = UMAP1, y = UMAP2, color = mean_cnv_detection)) +
      geom_point(alpha = 0.5, size = 0.5) +
      theme_bw() +
      scale_color_viridis_c() +
      facet_wrap(vars(reference_used)) + 
      labs(title = glue::glue("chr{chr}"))
  })
```

Looking at the UMAPs, I don't see any really obvious differences between CNV estimation and groups of cells. 
Let's look at the distribution of CNV estimation in cells that are called aneuploid and diploid by `CopyKAT`.

```{r}
# create faceted density plots showing estimation of CNV detection across each chr of interest 
# colored by aneuploid/diploid estimation 
ewings_chrs |> 
  purrr::map(\(chr){
    plot_df <- cnv_df |> 
      dplyr::filter(chromosome_name == chr)

    ggplot(plot_df, aes(x = mean_cnv_detection, color = copykat.pred)) +
      geom_density() +
      theme_bw() +
      facet_wrap(vars(reference_used)) + 
      labs(title = glue::glue("chr{chr}"))
  })
```

It looks like there is a slight increase in chr8 and chr12 indicating maybe a copy number gain for those chromosomes in aneuploid cells, but nothing striking.

>>>>

# join with cnv info
cnv_df <- cnv_df |> 
  dplyr::left_join(full_cnv_df, by = c("barcodes", "reference_used"))
```

Let's look at the distribution of CNV estimation in cells that are called aneuploid and diploid by `CopyKAT`.

```{r, fig.height=15, fig.width=10}
# create faceted density plots showing estimation of CNV detection across each chr of interest 
# colored by aneuploid/diploid estimation 
ggplot(cnv_df, aes(x = mean_cnv_detection, color = copykat.pred)) +
      geom_density() +
      theme_bw() +
      facet_grid(rows = vars(chromosome_name), 
                 cols = vars(reference_used))
```

It looks like the largest increases are in chr8 and chr21. 
Chr8 is one of the more common CNA's in Ewing's so this is a good sign.

Chr12 may show some increase, but Chr1 and Chr16 don't show any real differences.
